---
title: "Homework 6 -- Intro to Probability and Statistics"
author: "Raul Bassi"
output: pdf_document
---

# Instructions:

**Due:** 05/21 at 11:59PM.

**What am I expecting?** An R Markdown with the answers.

**Have fun!**

The Brazilian Ministry of Health keep records of the daily number of Coronavirus cases, by dates and municipalities. The dataset can be found here: (https://covid.saude.gov.br/). 

## Question 1

Take some time to familiarize yourself with the data. After that, please provide here a brief explanation of the variables and how the dataset is organized.

```{r include=FALSE}
library(dplyr)
```


Basically, the dataset is organized by State, number of cases confirmed, number of deaths, incidence by 100k habitants and death rate by 100k habitants and date of last update.



## Question 2

I saved the May 15 version of the dataset on the class GitHub, under the name `brazilcorona.csv`. Load the dataset in your computer and put a `head` of the dataset here.

```{r}
brcovid <- read.csv("https://raw.githubusercontent.com/umbertomig/intro-prob-stat-FGV/master/datasets/brazilcorona.csv", encoding="UTF-8")
head(brcovid)
```


## Question 3

Note that the dataset has the aggregated results by country, region, and state. Exclude all these aggregated results, keeping only the cases by the municipality. After that, exclude all cases before May 15. Put a `head` of the filtered dataset here. What is the dimension of the filtered dataset?

```{r}
covid19 <- filter(brcovid, municipio != "NA", data == "2020-05-15")
head(covid19)
```
3157 observables by 14 variables

## Question 4

Use the municipal population to compute the average infections rate (the number of infections divided by population times 100000) and mortality rate (the number of casualties divided by municipal population times 100000). For each state, which municipalities have the lowest and the highest infections and death rates? Create a new dataset with the average infections and mortality rates by municipality.
```{r}
covid19$mediadecasos <- (covid19$casosAcumulado/covid19$populacaoTCU2019)*100000
covid19$mediademortes <- (covid19$obitosAcumulado/covid19$populacaoTCU2019)*100000
covid_19 <- covid19[-c(1,4,5,6,7,9,12,13,14)]
top <- covid_19 %>% group_by(estado) %>% top_n(1, mediadecasos)
head(top, 27)
top1 <- covid_19 %>% group_by(estado) %>% top_n(-1, mediadecasos)
head(top1, 27)
top2 <- covid_19 %>% group_by(estado) %>% top_n(1, mediademortes)
head(top2, 27)
top3 <- covid_19 %>% filter(mediademortes != 0) %>% group_by(estado) %>% top_n(-1, mediademortes)
head(top3, 27)
corona <- covid19[c(3,5,15,16)]
```


## Question 5

In the dataset `brmayors.csv`, there are the party, gender, and schooling of the Brazilian mayors elected in 2016. I downloaded this dataset from the `CEPESP Data` repository (http://cepespdata.io/). Explore the variables making tables and/or barplots. Provide a brief interpretation of the results.

```{r}
brmayors <- read.csv("https://raw.githubusercontent.com/umbertomig/intro-prob-stat-FGV/master/datasets/brmayors.csv", encoding="UTF-8")
```
```{r}
table(brmayors$DESCRICAO_GRAU_INSTRUCAO)
table(brmayors$DESCRICAO_SEXO, brmayors$DESCRICAO_GRAU_INSTRUCAO)
barplot(table(brmayors$DESCRICAO_SEXO))
```
There are way more male mayors than females, most of them have college degrees, either men or women.

## Question 6

Merge this dataset with the mortality and infections rates dataset created in the previous question. Put a `head` of the new dataset.

```{r}
names(brmayors)[names(brmayors) == "COD_MUN_IBGE"] <- "codmun"
mycovid <- dplyr::full_join(corona, brmayors, by = "codmun")
head(mycovid)
```


## Question 7

Some experts showed that places ruled by women have lower mortality rates than areas ruled by men. Test this hypothesis using regression analysis. Provide a brief interpretation of your results.

```{r}
lm(mediademortes ~ DESCRICAO_SEXO, mycovid)
```
There is a small, but positive correlation on havin male mayors and higher death rates.

## Question 8

Note that most municipalities have no cases of the disease. What happens if you put zero in the places that had no cases, and run the regression in *question 6*? Provide a brief interpretation of your results.

```{r}
mycovid[is.na(mycovid)] <- 0
lm(mediademortes ~ DESCRICAO_SEXO, mycovid)
```
It is possible to understand that the correlation is positive in both, however, slightly smaller when considering all municipalities.

## Question 9

Create a variable that codes the ideological positions, based on the politician's parties. Use the three-way category: `left`, `center`, and `right`. Justify your choices. Provide a table with the frequency by each category. Run a regression for the new variable, to investigate whether ideological position influences the COVID casualties and infections. Provide a brief interpretation.

```{r}
l <- c("PCdoB", "PDT","PMN", "PSB", "PSOL","PT", "PV", "REDE", "PCO", "PSTU", "PCB", "PPS")
center <- c("SD", "PSDB", "PROS", "MDB")
mycovid$political_spectrum <- ifelse(mycovid$SIGLA_PARTIDO %in% l, 'Left',
                                     ifelse(mycovid$SIGLA_PARTIDO %in% center, 'Center', 'Right'))
table(mycovid$political_spectrum)
lm(mediadecasos ~ political_spectrum, mycovid)
lm(mediademortes ~ political_spectrum, mycovid)
  
```
To classify the parties, I utilized the "Congresso em Foco" spectrum created in 2019 ("https://congressoemfoco.uol.com.br/legislativo/direita-cresce-e-engole-o-centro-no-congresso-mais-fragmentado-da-historia/").

In the regression, it is possible to assess that LEft and Right parties have a negative correlation with deaths and with cases. Being with deaths close to zero, and with cases there is an increase in the negative tendence being Left smaller than right.


## Question 10

Using `ifelse` creates a variable `collegedegree`, which is `Yes` when the mayor has a college degree, and `No` otherwise. Run a regression for the new variable, to investigate whether education influences the COVID casualties and infections. Provide a brief interpretation.

```{r}
mycovid$collegedegree <- ifelse(mycovid$DESCRICAO_GRAU_INSTRUCAO == "SUPERIOR COMPLETO", 'Yes', 'No')
lm(mediadecasos ~ collegedegree, mycovid)
lm(mediademortes ~ collegedegree, mycovid)
```

Accordingly to the regression, there is a bigger correlation between having college degrees and the spread of the virus, however less correlation in deaths. That could be explained to the fact that bigger cities have more cases and most of them have mayors with degrees.