---
title: "Homework 8 -- Intro to Probability and Statistics"
author: "Raul Bassi"
output: pdf_document
---

# Instructions:

**Due:** 06/04 at 11:59PM.

**What am I expecting?** An R Markdown with the answers. Note: this one can be done in paper and pencil. If you want to do so, do the problem and take pictures of it. Put the pictures in a PDF and send it.

**Have fun!**

# Question 1

A factory produce valves, with 20\% chance of a given valve be broken. The valves are sold in boxes, containing ten valves in each box. If no broken valve is found, then they sell the box for \$10.00. With one broken valve, the box costs \$8.00. With two or three, the box is sold for \$6.00. More than three valves broken, they sell the box for \$2.00. What is the mean sales price for the boxes?

A box contain ten valves, each valve has 80\% of not being broken, therefore:
```{r}
mean(((0.8^10*1)*10.00)+((0.8^9*0.2*10)*8.00)+((0.8^8*0.2^2*45)*6.00)+((0.8^7*0.2^3*120)*6.00)+((0.8^6*0.2^4*210)*2.00)+((0.8^5*0.2^5*252)*2.00)+((0.8^4*0.2^6*210)*2.00)+((0.8^3*0.2^7*120)*2.00)+((0.8^2*0.2^8*45)*2.00)+((0.8^1*0.2^9*10)*2.00)+((0.2^10*1)*2.00))
```
Making a binomial distribution of $n\choose k$ being n = 10 and k values from 0 to 10 and summing all events of having 0 broken valves, 1 broken until 10 broken valves in a box, multiplied by their values. All in all it is a mean of $(p + (1-p))^{10}$

The answer is 6.482872

# Question 2

If a random variable $X$ has distribution:

|   x  |  -2 |  -1  |  0  |   1  |   2  |
|:----:|:---:|:----:|:---:|:----:|:----:|
| f(x) | 1/2 | 1/10 | 1/5 | 1/10 | 1/10 |
| F(x) | 1/2 |  3/5 | 4/5 | 9/10 |   1  |

Compute the mean, the variance, and the standard deviation of $X$.
```{r}
-2*1/2+-1*1/10+0*1/5+1*1/10+2*1/10
```


| x |$x-\mathbb{E}(x^2)$ |$x-\mathbb{E}(x^2)^2$|$x-\mathbb{E}(x^2)^2f(x)$|
|:-:|:------------------:|:-------------------:|:-----------------------:|
|-2 |   -2+0.8 = -1.2    |    -1.2^2 = 1.44    |    1.44 x 0.5 = 0.720   |
|-1 |   -1+0.8 = -0.2    |    -0.2^2 = 0.04    |    0.04 x 0.1 = 0.004   |
| 0 |    0+0.8 =  0.8    |     0.8^2 = 0.64    |    0.64 x 0.2 = 0.128   |
| 1 |    1+0.8 =  1.8    |     1.8^2 = 3.24    |    3.24 x 0.1 = 0.324   |
| 2 |    2+0.8 =  2.8    |     2.8^2 = 7.84    |    7.84 x 0.1 = 0.784   |

```{r}
sqrt(1.96)
```


The mean is -0,8

The variance is 1,96

The standard deviation is 1.4


# Question 3

If a random variable $X$ has distribution (note: it is the same as in the previous problem):

|   x  |  -2 |  -1  |  0  |   1  |   2  |
|:----:|:---:|:----:|:---:|:----:|:----:|
| f(x) | 1/2 | 1/10 | 1/5 | 1/10 | 1/10 |
| F(x) | 1/2 |  3/5 | 4/5 | 9/10 |   1  |

Compute the distribution of the following transformations:

* $X^2$?

|   x  |  0  |  1  |   4  |
|:----:|:---:|:---:|:----:|
| f(x) | 1/5 | 1/5 |  3/5 |
| F(x) | 1/5 | 2/5 |   1  |



* $3X$?

|   x  |  -6 |  -3  |  0  |   3  |   6  |
|:----:|:---:|:----:|:---:|:----:|:----:|
| f(x) | 1/2 | 1/10 | 1/5 | 1/10 | 1/10 |
| F(x) | 1/2 |  3/5 | 4/5 | 9/10 |   1  |

# Question 4

Prove the following statements:

* Let two constants, $a \in \mathbb{R}$ and $b \in \mathbb{R}$, and a discrete random variable $X$. Prove that $\mathbb{E}(aX+b) = a\mathbb{E}(X) + b$.

$$\mathbb{E}(aX+b) = \sum x(ax+b)p(x)$$
$$\mathbb{E}(aX+b) = \sum x .a.x.p(x)+\sum x .b.p(x)$$
$$\mathbb{E}(aX+b) = a(\sum x .x.p(x))+b(\sum x.p(x))$$
$$1 =\sum x p(x) = \mathbb{E}(x)$$
$$\mathbb{E}(aX+b) = a\mathbb{E}(x)+b$$

* Let two constants, $a \in \mathbb{R}$ and $b \in \mathbb{R}$, and a discrete random variable $X$. Prove that $\mathbb{V}(aX+b) = a^2\mathbb{V}(X)$.

$$\mathbb{V}(aX+b)=\mathbb{E}((ax+b)^2) - (\mathbb{E}(aX+b))^2$$
$$=\mathbb{E}(ax^2 + 2axb + b^2) - (\mathbb{E}(aX+b))(\mathbb{E}(aX+b))$$
$$=a^2\mathbb{E}(X^2) + 2ab\mathbb{E}(X) + b^2 - a^2(\mathbb{E}(X))^2 - 2ab\mathbb{E}(x)-b^2$$
$$=a^2\mathbb{E}(X^2) - a^2(\mathbb{E}(X))^2$$
$$a^2[\mathbb{E}(X^2) - (\mathbb{E}(X))^2]$$
$$\mathbb{E}(X^2) - (\mathbb{E}(X))^2 = \mathbb{V}(X)$$
$$\mathbb{V}(aX+b) = a^2\mathbb{V}(X)$$





* Let $a \in \mathbb{R}$ a constant. Prove that $\mathbb{E}(a) = a$ and that $\mathbb{V}(a) = 0$.
$f(x_i) = a$
* $\mathbb{E}(x) = \sum x_if(x_i)$
* $\mathbb{E}(a) = \sum x_if(x_i)$
* $f(x_i) = a + 0*x_i$
* Therefore, it does not matter what n value multiplies Xi bc it will always be 0 + a. Thus:
* $\mathbb{E}(a) = a$

Knowing this and given that:

$$\mathbb{E}(X^2) - (\mathbb{E}(X))^2 = \mathbb{V}(X)$$

$$\mathbb{V}(X) = (a^2 - a^2) = 0$$
