---
title: "Midterm -- Stats for IR"
author: "Raul Bassi"
date: "Due 4/23/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

**Instructions:**

1. This is a 10-days take-home exam.
2. This is an individual exam. You should not collaborate.
3. I will hold office hours Thursday, during class time.
4. In the office hours, you may not ask questions specifically about the midterm. Please rephrase the issues in terms of similar problems.
5. If it does not knit, the grade is **zero**.
6. Good luck!

## Question 1

Thomas Fujiwara wrote the [following tweet](https://twitter.com/thomasfujiwara/status/1249817958874001412?s=20) to analyze the COVID-19 evolution. You should replicate the same graph, not for the Brazilian States, but countries. After this, you should tell us about the COVID-19 trends in the world. In which countries the COVID is going down? In which countries the COVID is going up? Do two separate analysis for the:

1. Evolution of infections
2. Evolution of casualties

**Hint:** recycle the code you wrote in the last problem set.

1)
```{r}
download.file("https://github.com/RamiKrispin/coronavirus/raw/master/data/coronavirus.rda", "cv")
load("cv")
library(dplyr)
at <- coronavirus %>%group_by(country, date, type) %>% summarise(total_cases = sum(cases)) %>% arrange(-total_cases)
aw <- coronavirus %>%group_by(country, type) %>% summarise(total_cases = sum(cases)) %>% arrange(-total_cases)
at1 <- filter(at, type == "confirmed")
at2 <- filter(at1, date == max(at1$date))
## TOP 10 INFECTIONS 
at2 %>% head(10)
## TOP 10 LEAST INFECTIONS
at2 %>% tail(10)
```
```{r}
gg <- coronavirus %>%group_by (country, date, type) %>% summarise(total_cases = sum(cases)) %>% arrange(-total_cases)
infection <- filter(gg, type == 'confirmed')
target <- c("US", "France", "Spain", "United Kingdom", "Italy", "Brazil", "China")
t10 <- filter(infection, country %in% target)
t11 = t10[c(-3)] 
```
```{r}
library("reshape2")

"Log Scale"
ggplot(data=t11, aes(x=date, y=log(total_cases), group = country, colour = country)) +
    geom_smooth(se = FALSE) +
  labs(title = "Infections by date(log version)", 
       x = "dates",
       y = "Cases registered(log)")
"No log scale"
ggplot(data=t11, aes(x=date, y=total_cases, group = country, colour = country)) +
    geom_smooth(se = FALSE) +
    geom_point( size=1, shape=21, fill="white") +
  labs(title = "Infections by date", 
       x = "dates",
       y = "Cases registered")
```


2)
```{r}
at3 <- filter(at, type == "death")
at4 <- filter(at3, date == max(at1$date))
## TOP 10 DEATHS 
at4 %>% head(10)
## TOP 10 LEAST DEATHS
at4 %>% tail(10)
```

```{r}
death <- filter(gg, type == 'death')
target2 <- c("US", "France", "Spain", "United Kingdom", "Italy", "Brazil", "China")
t12 <- filter(death, country %in% target2)
t13 = t12[c(-3)]
"Log scale"
ggplot(data=t13, aes(x=date, y=total_cases, group = country, colour = country)) +
    geom_smooth(se = FALSE) +
  labs(title = "Deaths by date(log version)", 
       x = "dates",
       y = "Deaths registered(log)") + scale_y_log10()
"No log scale"
ggplot(data=t13, aes(x=date, y=total_cases, group = country, colour = country)) +
    geom_smooth(se = FALSE) +
    geom_point( size=1, shape=21, fill="white") +
  labs(title = "Deaths by date", 
       x = "dates",
       y = "Deaths registered")
```
## Question 2

Let the following vector:

```{r}
ages <- rbinom(100, 50, .2)+18
```

Compute:

1. The mean of the vector.
2. The median of the vector.
3. The quartiles.
4. The 1-percentile and the 99-percentile.
5. Make a histogram.
6. Make a box-plot.

```{r}

mean(ages)

median(ages)

quantile(ages)

quantile(ages, c(.01, .99))

hist(ages, main="Histogram", xlab="Age", ylab="Frequency", col.main="orange", col.lab="blue")

boxplot(ages, main="Boxplot", ylab="Age", col.main="yellow", col.lab="gray")
```
# Question 3

In the following dataset, we have an experiment where researchers send resumes for jobs, randomizing the names of the applicants. In the US, the names of applicants are correlated with race, and this allows firms to select racially based on the person's names.

The paper that this problem came from is the following: Marianne Bertrand and Sendhil Mullainathan (2004) “Are Emily and Greg more employable than Lakisha and Jamal? A field experiment on labor market discrimination.” American Economic Review, vol. 94, no. 4, pp. 991–1013. (https://www.aeaweb.org/articles?id=10.1257/0002828042002561)

The dataset is the following:

```{r}
resume <- read.csv('https://raw.githubusercontent.com/umbertomig/intro-prob-stat-FGV/master/datasets/resume.csv')
head(resume)
```

If has the following variables:

* **firstname:** First name in the resume
* **sex:** Gender of the applicant
* **race:** Race of the applicant
* **call:** Whether a callback was made (call=1) or not (call=0)

Answer the following questions:

1. What is the overall callback rate?
2. What is the black callback rate?
3. What is the white callback rate?
4. What is the Black-Female callback rate?
5. What is the White-Female callback rate?
6. What is the callback rate by the first name? Which first name receives the highest and the lowest callback rates?
7. Make a mosaicplot of the callback rate by race.
8. Make a mosaicplot of the callback rate by gender.
```{r}
table(resume$call)
392/nrow(resume)
```
1-
4478 were not called back 
392 were called back
8% of all were called back

```{r}
table(resume$call, resume$race)

157/(2278 +157) 
235/(2200 +235)
```
2-
2278 blacks were not called back
392  blacks were called back
6,4% of blacks were called back

3-
2200 whites were not called back
235 whites were called back
9,6% of whites were called back

```{r}
table(resume$sex,resume$race,resume$call)
125/(1761 + 125)
184/(1676+184)
```
4 - BLACK FEMALE
RATE 6,6%
Not called back: 1761
Called back: 125

5 - WHITE FEMALE
RATE 9,8%
Not called back: 1676
Called back: 184

6-
```{r}
summary_RE <- resume %>% group_by(firstname) %>%
  summarise(call = sum(call)) %>%
  arrange(-call)
r1 = resume[-c(2,3,4)]
r2 <- as.data.frame(table(r1))
names(r2)[names(r2) == "r1"] <- "firstname"
r3 <- total <- merge(r2,summary_RE,by="firstname")
r3$callback_rate <- (r3$call)/(r3$Freq)
summary_RE %>% head(1)
summary_RE %>% tail(1)
r3
```
Kristen is the most called back name
Rasheed is the least called back name

7-
```{r}
mosaicplot(table(resume$call, resume$race), main = "CALL X RACE", color = "lightblue")
```
8- 
```{r}
mosaicplot(table(resume$call, resume$sex), main = "CALL X GENDER", color = "orange")
```


## Question 4

1. What is the fundamental problem of the causal inference? Provide an example.
2. One researcher wants to study tax conservatism among the rich. She selects lottery ticket winners to run her study, and she claims that the biggest the ticket the person won, the more the person is against income taxes. Should we believe this study? What are the strengths of this study? What are the weaknesses of this study?
3. Suppose you want to replicate the study in **question 3** in Brazil. What would you do? What are the design choices that you would make? What are the potential problems?

1) The fundamental problem of the causal inference is that it is impossible to analyse the effects of treatment and non-treatment on the same unit. A example is the use of medicine to treat headache. The person either takes or not takes de medicine, becausse if it takes it is impossible to analyse what would be the effects on the contrary in the same person.

2) Yes, because considering two persons, Person A and Person B. If Person A won in the lottery $1,000,000 while the person B won $10,000,000. Both pay the same tax rate of 15%. Person A would be pay $150,000 while the person B would pay $1,500,000. Proportionally, is the same amount of money, however there is a psychological effect. Whixh is worst giving the government $150,000 or $1,500,000?
The strenght of this study is that it shows how lottery winners behave, however it fails because takes only a small part of the small group of the riches, only those who won the lottery. In this sense, I do not think it is possible to derivate the behave of people who won the lottery with people such as Jeff Bezos or Bill Gates.

3)To replicate the study in Brazil, firstly, I would have to find another feature that differenciate blacks and whites here. Because first names does not mean anything here. The randomization choice is a must, probably I would adopt data such as how many black people are employed in different sectors of society and use this feature to see the tendence of such sectors. One big problem is that unemployement rate in Brazil is very high. Thus, I think sending fake resumes would not be something very ethical to be done. Considering some of them have chances of being selected and taking place on the interview of a real person.